---
title: "Author Classes And Methods That Extend The Ready4 Framework"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Author Classes And Methods That Extend The Ready4 Framework}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE}
library(ready4class)
library(ready4show)
library(ready4use)
```

## Motivation

The [ready4 framework](https://www.ready4-dev.com/) uses [object oriented programming (OOP)](https://ready4-dev.github.io/ready4/articles/V_03.html) to implement [modular approaches to computational models of mental health systems](https://ready4-dev.github.io/ready4/articles/V_01.html). That means that a standardised approach to developing modules (S4 classes) and sub-modules (S3 classes) is required. ready4class provides the tools to implement this workflow.

## Workflow

`ready4class` is an extension of the workflow tools for authoring ready4 framework R package functions that are described briefly in [this vignette from the ready4fun package](https://ready4-dev.github.io/ready4fun/articles/V_01.html). The tools included in `ready4class` are intended for use as part of [ready4pack R package authoring workflow](https://ready4-dev.github.io/ready4pack/articles/V_01.html).

The two classes exported as part of `ready4class` that are used in the `ready4pack` workflow are `ready4class_constructor` and `ready4class_pt_lup`. `ready4class_constructor` is a tibble based class in which the properties of the classes to be authored are declared. Instances of `ready4class_constructor` are most efficiently created using the `make_pt_ready4class_constructor` function. Examples of how to use `ready4class_constructor` and `ready4class_pt_lup` are referenced in the [a vignette from the ready4pack package](https://ready4-dev.github.io/ready4pack/articles/V_01.html)

`ready4class_pt_lup` is another tibble based class, which contains metadata on the prototypes of classes that can be used as sub-components of other classes (for example a tibble based class can be used as a slot in an S4 class). When authoring ready4 R packages, you will create a `ready4class_pt_lup` and store it in an online repository that you have write permissions to. As you create new classes using `ready4class` tools, your `ready4class_pt_lup` object will be updated so that these classes can be easily incorporated into any other classes you make in future. An example of an `ready4class_pt_lup` object is reproduced below.

```{r ptlup, tab.cap='Class Prototypes Lookup Table', tab.id = 'ptlup', results = "asis"}
x <- ready4use::Ready4useRepos(gh_repo_1L_chr = "ready4-dev/ready4",
                               gh_tag_1L_chr = "Documentation_0.0") %>%
  ingest(fls_to_ingest_chr = "prototype_lup",
         metadata_1L_lgl = F) 
exhibit(x)
```

## Future documentation

It should be noted that some `ready4class` methods require files of a standardised format to be saved in specific sub-directories of the package `data-raw` directory. Detailed instructions on how to prepare these files are not yet available, but will be outlined in documentation to be released in 2022.

