---
title: "Author Classes And Methods That Extend The Ready4 Framework"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Author Classes And Methods That Extend The Ready4 Framework}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE}
library(ready4class)
library(ready4show)
library(ready4use)
```

The [ready4 framework](https://www.ready4-dev.com/) for open and modular mental health systems models is implemented using object oriented programming. That means that the framework largely consists of classes (representations of data structures useful for modelling mental health systems) and methods (algorithms that can be applied to these data-structures to generate insight useful for policy-making).

## Rationale

There are numerous existing third party R packages (`devtools`, `methods` and others) that are useful for authoring and documenting classes and methods. The tools contained in these packages present developers with a great degree of flexibility in how to approach object oriented programming within R. However, the more developers exercise this flexibility, the less standardised the output.  `ready4class` has been developed to help ensure that developers extending the `ready4 framework` apply consistent naming, representation and documenting conventions when authoring classes and methods for R packages.

## Workflow

`ready4class` is an extension of the workflow tools for authoring ready4 framework R package functions that are described briefly in [this vignette from the ready4fun package](https://ready4-dev.github.io/ready4fun/articles/V_01.html). The tools included in `ready4class` are intended for use as part of [ready4pack R package authoring workflow](https://ready4-dev.github.io/ready4pack/articles/V_01.html).

## Contents

The two classes exported as part of `ready4class` that are used in the `ready4pack` workflow are `ready4class_constructor` and `ready4class_pt_lup`. `ready4class_constructor` is a tibble based class in which the properties of the classes to be authored are declared. Instances of `ready4class_constructor` are most efficiently created using the `make_pt_ready4class_constructor` function. Examples of how to use `ready4class_constructor` and `ready4class_pt_lup` are referenced in the [a vignette from the ready4pack package](https://ready4-dev.github.io/ready4pack/articles/V_01.html)

`ready4class_pt_lup` is another tibble based class, which contains metadata on the prototypes of classes that can be used as sub-components of other classes (for example a tibble based class can be used as a slot in an S4 class). When authoring ready4 R packages, you will create a `ready4class_pt_lup` and store it in an online repository that you have write permissions to. As you create new classes using `ready4class` tools, your `ready4class_pt_lup` object will be updated so that these classes can be easily incorporated into any other classes you make in future. An example of an `ready4class_pt_lup` object is reproduced below.

```{r ptlup, tab.cap='Class Prototypes Lookup Table', tab.id = 'ptlup', results = "asis"}
x <- ready4use::Ready4useRepos(gh_repo_1L_chr = "ready4-dev/ready4",
                               gh_tag_1L_chr = "Documentation_0.0") %>%
  ingest(fls_to_ingest_chr = "prototype_lup",
         metadata_1L_lgl = F) 
exhibit(x)
```

## Future documentation

It should be noted that some `ready4class` methods require files of a standardised format to be saved in specific sub-directories of the package `data-raw` directory. Detailed instructions on how to prepare these files are not yet available, but will be outlined in documentation to be released in 2022.

