---
title: "Author Ready4 Framework Modules"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Author Ready4 Framework Modules}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE}
library(ready4class)
library(ready4show)
library(ready4use)
```

## Motivation

The [ready4 framework](https://www.ready4-dev.com/) uses [object oriented programming (OOP)](https://ready4-dev.github.io/ready4/articles/V_03.html) to implement [modular approaches to computational models of mental health systems](https://ready4-dev.github.io/ready4/articles/V_01.html). That means that a standardised approach to developing modules (S4 classes) and sub-modules (S3 classes) is required. ready4class provides the tools to implement this workflow.

## Workflow

### Prototyes, constructor and manifest
The main classes exported as part of `ready4class` are `readyclass_manifest` and `ready4class_constructor`. 
`ready4class_pt_lup` is a tibble based [ready4 Framework sub-module](https://ready4-dev.github.io/ready4/articles/V_01.html), which contains metadata on the prototypes of classes that can be used as sub-components of ready4 Framework modules and sub-modules (for example a tibble based class can be used as a slot in an S4 class). When authoring ready4 R packages, you will create a `ready4class_pt_lup` instance and store it in an online repository that you have write permissions to. As you create new ready4 Framework modules and sub-modules using `ready4class` tools, your `ready4class_pt_lup` object will be updated so that these classes can be made available to any future modules or sub-modules that you author. The first few entries in an example `ready4class_pt_lup` sub-module are reproduced below.

```{r ptlup, tab.cap='Class Prototypes Lookup Table', tab.id = 'ptlup', results = "asis"}
x <- ready4use::Ready4useRepos(gh_repo_1L_chr = "ready4-dev/ready4",
                               gh_tag_1L_chr = "Documentation_0.0") %>%
  ingest(fls_to_ingest_chr = "prototype_lup",
         metadata_1L_lgl = F) 
x %>%
  head() %>%
  exhibit()
```

`ready4class_constructor` is another tibble based ready4 Framework sub-module that summarises the desired features of the ready4 Framework modules and sub-modules that you are authoring. An instance of `ready4class_constructor` is combined with a [`ready4fun_manifest` sub-module](https://ready4-dev.github.io/ready4fun/articles/V_01.html) to create a `ready4class_manifest` sub-module. Instances of `ready4class_constructor` are most efficiently created using the `make_pt_ready4class_constructor` function.

```{r warning=FALSE}
y <- make_pt_ready4class_manifest(ready4fun::ready4fun_manifest(),
                                  constructor_r3 = ready4class_constructor()) %>%
  ready4class_manifest()
```

### Typical use
The most important method included in `ready4class` is the `author` method for the `ready4class_manifest` sub-module, that enhances [the `author` method defined for the `ready4fun_manifest`](https://ready4-dev.github.io/ready4fun/articles/V_01.html) so that consistently documented R package classes are also generated.

```{r eval = F}
## Not run
author(y)
```

### Examples
`ready4class` sub-modules and methods are not intended for independent use, but instead should be deployed as part of [ready4pack R package authoring workflow](https://ready4-dev.github.io/ready4pack/articles/V_01.html).

## Future documentation

It should be noted that some `ready4class` methods require files of a standardised format to be saved in specific sub-directories of the package `data-raw` directory. Detailed instructions on how to prepare these files are not yet available, but will be outlined in documentation to be released in 2022.
